---
title: "R2D2 Prior for Arbitrary Link Function (Probit Model)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{probit_example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

In this example we will demonstrate how to use the "arbitrary link" functionality of the R2D2 prior with the probit model. Let $Y_i\in\{0,1\}$ be the response and $X_i$ be the $(p=10)$ fixed effect covariates for $i=1,\dots,100$. Then
$$
  Y_i \sim\mathsf{Bernoulli}(\Phi(\eta_i))
$$
where
$$
  \eta_i = \beta_0 + X_i\beta
$$
and $\Phi(\cdot)$ is the cumulative distribution function (CDF) for a standard normal random variable.

## Generate Data

```{r}

set.seed(12345)

n <- 100
p <- 10

beta0 <- 0.25
beta <- rnorm(p)

X <- matrix(rnorm(n*p), ncol=p, nrow=n)
eta <- beta0 + X%*%beta

Y <- rbinom(n, 1, pnorm(eta))

```


#### Priors

$$
  \beta_0\sim\mathsf{Normal}(\mu_0,\tau_0^2),\ \beta_j\sim\mathsf{Normal}(0,\phi_j W), (\phi_1,\dots,\phi_p)\sim\mathsf{Dirichlet}(\xi_1,\dots,\xi_p),\
  W\sim\mbox{GBP}(a^*,b^*,c^*,d^*)
$$
where $\mu_0=0,\tau_0^2=3$ and $\xi_j=1$ for $j=1,\dots,p=10$.

### Compute $(a^*,b^*,c^*,d^*)$

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(r2d2glmm)
```

```{r}
a=1
b=1

mu <- function(x){pnorm(x)}
sigma2 <- function(x){pnorm(x)*(1-pnorm(x))}


params <- WGBP(a, b, b0=qnorm(mean(Y)), link="Arbitrary", mu=mu, sigma2=sigma2)
params
```



### Fit Model in JAGS

```{r}
library(rjags)

  data = list(Y=Y, X = X, n=n, p=p, params=params)
  
  model_string <- textConnection("model{
                               
  ##Likelihood

 for(i in 1:n){
    Y[i] ~ dbern(mu[i])  
    probit(mu[i]) <- beta0 + inprod( X[i, ], beta[])
 }


                              
##Priors


  V ~ dbeta(params[1], params[2])
  W <- params[4]*(V/(1-V))^(1/params[3])
  
  for(i in 1:p){delta[i] <- 1}
  dd ~ ddirch(delta)
  
  beta0 ~ dnorm(0, 1/3)
  
  for(j in 1:p){
  beta[j] ~ dnorm(0, 1/(W*dd[j]))
  }
  
        
                                 }")
  
  inits <- list(beta=rnorm(p,0,1))
  model <- jags.model(model_string,data = data, inits=inits, n.chains=1)
  
  update(model, 5000)
  
  params <- c("beta", "beta0", "W", "dd")
  samples <- coda.samples(model,
                          variable.names=params,
                          n.iter=10000)

```


## Check convergence

```{r, fig.width=7}
plot(samples[[1]][,1]) # W
plot(samples[[1]][,3]) # beta_2
plot(samples[[1]][,5]) # beta_4
plot(samples[[1]][,1+p+2]) # phi_1

```


## Plot results

```{r, fig.width=7}
library(dplyr)
library(ggplot2)

bayes_R2 <- function(X, p, samples){
  
  b0s <- as.matrix(samples[[1]][,1+p+1])
  betas <- as.matrix(samples[[1]][,(2):(1+p)])

  etas <- betas%*%base::t(X) 
  
  etas <- sweep(etas, 1, b0s, "+")
  
  y_pred <- pnorm(etas)
  y_err  <- pnorm(etas)*(1-pnorm(etas))
  
  M <- apply(y_pred,1,var)
  V <- apply(y_err,1,mean)
  return(M/(M+V))
  
}


r2.df <- tibble(r2=bayes_R2(X,p,samples))

ggplot(data = r2.df, aes(x=r2))+
  geom_density()+
  xlab("R2")+
  ylab("Density")+
  ggtitle("Bayesian Posterior R2")+
  theme_minimal()



```









